%------------------------------------------------------------------------------
\chapter{Acknowledgements}
\label{sec:ack}
%------------------------------------------------------------------------------

During the realization of this thesis, there has been a group of people without whom it would not have been possible. Due to this, they deserve to be thanked and mentioned here. I would like to start with Prof. Dr. Martin Weitz, for giving me the opportunity of carrying out a thesis with his research group and helping me to achieve it in every step of the way.

The next person I would like to thank is Roberto R\"oll. His work inside and outside the erbium laboratory has been crucial for the achievement of results obtained here. He is a person that has taught me a lot. From how to run such a complex experiment like this one, to more professional and personal skills, like how to work in team or how to keep a good attitude even when things were not working the way they should (which was in most cases).

Another one that must be mentioned is Dr. Arif Warsi Laskar. For his hard work and patience with my many questions and doubts, this thesis would also not have been possible without him, thank you. I would also like to thank David Helten, for his inspiring passion, help and good mood at every time we had to work hard. Moreover, I also need to thank the whole group of Quantum optics for making such a great environment and willingness to help at any time.

Also, I would like to thank both my brother and parents, for their emotional and economical support during all these years. Without it, I would not have been able to reach where I am now. Lastly, I want to thank Carrie for supporting me at the end of everyday. To everyone who has been by my side, thank you.



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../Template/Thesis"
%%% End: 
